An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-22 Revision/response.tex

\noindent
\textbf{Response:} 
That's a great suggestion. In order to figure out if the server setting affects the relative ordering of methods, we conducted additional experiments on a server equipped with two Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset.
The results show that the performance differences across algorithms remain small, and their relative rankings do not change significantly across servers.
We did observe slightly lower QPS on this server compared to our original one, which is mainly due to the lower CPU frequency of the Xeon 6330 processors. Please see section 4.2.4.

\textcolor{blue}{"To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on a server equipped with two Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure 12(b). Compared with the experimental results on the original server (as shown in Figure 4d), the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency of the Xeon 6330 compared to the original server, the QPS values are slightly lower across all methods."}

%%%%%%%%%%%%%%%%%%%%%% R6 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{In Table 1, please clarify which methods are disk-based, and which methods are memory-based.
			For disk-based methods, it is also meaningful to report the cost breakdown (e.g., in terms of I/O time and CPU time).}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. Currently, except for FilteredDiskANN, other hybrid query methods are memory-based, so we added the feature of whether to support disk for each algorithm in Table 1. At the same time, we also compared the search performance difference of FilteredDiskANN on memory and SSD using real data sets. The results are shown in Figure\ref{fig:recall-latency}:
\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/recall_latency_all.png}
	\caption{}
	\label{fig:recall-latency}
\end{figure}

We conducted additional experiments on the real-world YT8M dataset, comparing three variants of DiskANN.
(1) SSD-based FilteredDiskANN (original disk-based method);
(2) FilteredVamana (memory-based method);
(3) StitchedVamana (memory-based method).

%\begin{itemize}
%	\item SSD-based FilteredDiskANN (original disk-based method)
%	\item FilteredVamana (memory-based method)
%	\item StitchedVamana (memory-based method)
%\end{itemize}

@@

All experiments were run with 32 search threads for fair comparison. The following figure shows the mean latency (log-scale, left y-axis) and I/O vs CPU proportion (right y-axis) across different Recall\@10 levels:

\textbf{ Key Observations:}
Latency-Recall Tradeoff:
As expected, SSD-based FilteredDiskANN exhibits significantly higher latency than in-memory methods, especially at higher recall levels. When Recall@10 reaches 0.96, its latency increases super-linearly to over 100 ms.

@@

Cost Breakdown of DiskANN:The I/O proportion consistently dominates, maintaining 80\%–85\% of the total latency even at different recall levels.
The CPU proportion remains low (10–15\%), indicating that I/O is the major bottleneck in disk-based search pipelines.
This clearly shows that improving disk read efficiency or reducing I/O volume is critical for performance optimization.

\textbf{In-memory Methods:}
FilteredVamana achieves a good balance, staying under 30 ms while achieving Recall@10 > 0.97.
While StitchedVamana achieves extremely low latency (sub-5 ms) due to its stitched multi-label index structure, its fixed edge connectivity may limit effectiveness under highly uncorrelated label distributions. Nevertheless, it consistently supports 90\%+ Recall@10 across diverse datasets, as shown in our evaluations.

\textbf{Note: }Due to space limitations and the fact that only one method supports disk-based execution, we did not include the above experiments and analysis in the revised paper.

### 2025-07-22 Revision/response.tex

\textbf{In-memory Methods:}
FilteredVamana achieves a good balance, staying under 30 ms while achieving Recall@10 > 0.97.
While StitchedVamana achieves extremely low latency (sub-5 ms) due to its stitched multi-label index structure, its fixed edge connectivity may limit effectiveness under highly uncorrelated label distributions. Nevertheless, it consistently supports 90\%+ Recall@10 across diverse datasets, as shown in our evaluations.

\textbf{Note: }Due to space limitations and the fact that only one method supports disk-based execution, we did not include the above experiments and analysis in the revised paper.

%%%%%%%%%%%%%%%%%%%%%% R7 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Report the peak memory usage of query processing.}
	\end{metarevision}
}

Thank you for your suggestion. We are deeply aware of this shortcoming, so we provide the search peak memory usage of each algorithm in the time and space overhead section and carefully analyze the reasons.
\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/searchMem/label_memory_comparison.pdf}
	\caption{attribute filtering search peak memory}
	\label{fig:attribute filtering search peak memory}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/searchMem/range_memory_comparison.pdf}
	\caption{range filtering search peak memory}
	\label{fig:range filtering search peak memory}
\end{figure}

For attribute filtering, please see section 4.2.1.

\textcolor{blue}{"As for the search peak memory usage, we can find from the Figure ~\ref{fig:attribute filtering search peak memory} that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the peak memory usage of database is comparatively lower."}

For range filtering, please see section 4.3.1.

\textcolor{blue}{"As for the Search peak memory usage, we can find from the Figure ~\ref{fig:range filtering search peak memory} that it is basically consistent with the Build peak memory usage trend. The algorithm needs to load the complete index into memory during search. Therefore, algorithms with high building overhead, such as WinFilter, also have high search overhead; algorithms with low building overhead, such as Faiss, also have low search overhead. For Milvus, search does not require full index loading and there is additional metadata overhead when building, so the performance is different from the trend."}

### 2025-07-22 Revision/response.tex

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/attribute_selectivity.png}
	\caption{}
	\label{fig:attribute_selectivity}
\end{figure}

\textcolor{blue}{"The proximity of the four curves (representing different attribute distributions) in the figure reflects the algorithm's sensitivity to distributional variations. When the curves converge closely, the algorithm demonstrates strong resilience against attribute distribution shifts. Conversely, greater divergence among the curves indicates higher distributional dependence. Accordingly, the "Strong" label signifies robust performance consistency across varying distributions, while "Weak" denotes pronounced performance fluctuations under distributional changes."}

%%%%%%%%%%%%%%%%%%%%%% R2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Run experiments with multi-modal queries (e.g., combining text and images) or more complex query types involving both attribute and range filters.}
	\end{metarevision}
}

\noindent
\textbf{Response:} We believe this is an excellent suggestion. In fact, in the initial version of our paper, we used the text-to-image dataset that has been widely adopted by existing multimodal ANN algorithms. In this dataset, the base vectors are images, while the query set consists of texts, resulting in a modality gap between the two. Achieving good performance on such a dataset often requires the algorithm to demonstrate stronger robustness. \textbf{(Please refer to the last paragraph of Section 4.2.3 and the green highlighted part in Section 4.3.3.)}

To further enrich the evaluation of complex multimodal queries, we have added new experiments in the Multi-modal Queries sections of Section 4.2.4 and Section 4.3.4. Specifically, we constructed a more challenging multimodal dataset by selecting 500,000 text vectors and 500,000 image vectors as the base set, and 5,000 text vectors and 5,000 image vectors as the query set. We then performed both attribute filtering and range queries over this dataset, combining text and image queries to further analyze the robustness of the algorithms. The corresponding experimental results are presented in Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we did not consider this type of complex query in our experiments but instead studied the two query types separately. \textbf{(Please refer to the blue highlighted parts in Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.)
}

### 2025-07-22 VLDB2026/main.tex
1
change multimodal
修改多模态相关部分回复格式

	Most algorithms struggle to achieve high recall on GloVe, because it has the highest LID. Compare to other algorithms, the 4 algorithms (UNG, Faiss, CAPS, and Milvus) manage to maintain relatively high recall, demonstrating better adaptability to complex datasets. Notably, although vanilla Faiss performs worse on SIFT1M compared to Enron dataset, Faiss+HQI\_Batch outperforms on SIFT1M. Because batch optimization in HQI reduces query overhead and increases throughput.

	Most algorithms exhibit the worst performance on the OOD dataset Text2Image. However, ACORN, UNG, and FilteredVamana show relatively stable performance on Text2Image. Their performance drops less compared to in-distribution (ID) datasets such as GIST1M and GloVe, indicating better adaptability to OOD queries.}

@@

	\sout{To evaluate the robustness of the RF-ANNS algorithms in different scenarios, we add an OOD dataset in the experiments. As shown in Figure~\ref{fig:exp_8_3}, most methods perform worst on the Text2Image dataset. This is because OOD queries make the $k$-nearest neighbor distribution more sparse. As a result, the search needs to visit more nodes, which leads to an exponential growth in the search space and a sharp drop in efficiency.

	WinFilter shows the worst performance on the WIT dataset. This suggests that its performance is sensitive to the data dimensionality. ACORN has large performance differences across datasets, indicating weak robustness.

>>>>

	Most algorithms struggle to achieve high recall on GloVe, because it has the highest LID. Compare to other algorithms, the 4 algorithms (UNG, Faiss, CAPS, and Milvus) manage to maintain relatively high recall, demonstrating better adaptability to complex datasets. Notably, although vanilla Faiss performs worse on SIFT1M compared to Enron dataset, Faiss+HQI\_Batch outperforms on SIFT1M. Because batch optimization in HQI reduces query overhead and increases throughput.	}

	\textcolor{green}{Most algorithms exhibit the worst performance on the OOD dataset Text2Image. However, ACORN, UNG, and FilteredVamana show relatively stable performance on Text2Image. Their performance drops less compared to in-distribution (ID) datasets such as GIST1M and GloVe, indicating better adaptability to OOD queries.}

@@

	\textcolor{green}{To evaluate the robustness of the RF-ANNS algorithms in different scenarios, we add an OOD dataset in the experiments. As shown in Figure~\ref{fig:exp_8_3}, most methods perform worst on the Text2Image dataset. This is because OOD queries make the $k$-nearest neighbor distribution more sparse. As a result, the search needs to visit more nodes, which leads to an exponential growth in the search space and a sharp drop in efficiency.}

	\sout{WinFilter shows the worst performance on the WIT dataset. This suggests that its performance is sensitive to the data dimensionality. ACORN has large performance differences across datasets, indicating weak robustness.
